Age-related changes in plant growth and traits

### 2016-03-17 Accessory_figs_20160218.Rmd
accessory cost work

#seed count, scaled to plant size, decreases 1:1 with seed size, while bud count decreases less steeply; indication that seed count is what is actually "programmed in"
```{r}

win.metafile("figs2/Accessory_Costs/Figure_cost_of_flowers_vs_seed_size.wmf", height=5, width=15)
plot

mod1 <- sma(scaled_seed_count~seed_size,data=subset(SummaryInd,scaled_seed_count!=0),log="xy",slope=-1,method="OLS")
mod2 <- sma(scaled_bud_count~seed_size,data=subset(SummaryInd,scaled_bud_count!=0),log="xy",slope=-1,method="OLS")
mod3 <- sma(scaled_failure_count~seed_size,data=subset(SummaryInd,scaled_failure_count!=0),log="xy",slope=-1,method="OLS")
sma_abline(mod1)
abline(mod2$coef[[1]][1,1],mod2$coef[[1]][2,1],col="blue")
abline(mod3$coef[[1]][1,1],mod3$coef[[1]][2,1],col="red")
legend("topright",legend=c("seed count","total bud count","failed flower count"),col=c("black","blue","red"),pch=c(16,21,8),bty="n")

#Neither within nor across species, does the proportion of reproductive energy spent on "success" correlate with total reproductive investment
```{r}
par(mfrow=c(1,1), cex=1, omi=c(.1,.1,.1,.05), mai=c(.8,.8,.1,.05))

@@

write.csv(results, file="figs2/Accessory_Costs/Table02correlations_with_repro_inv.csv",row.names=FALSE)
```

#Figure 1. Correlations of various reproductive investment categories against plant weight
```{r}
win.metafile("figs2/Accessory_Costs/Figure01_correlations_with_plant_weight.wmf", height=5, width=10)
plot

@@

write.csv(results, file="figs2/Accessory_Costs/Table02extra_propagule_invest_vs_repro_inv.csv",row.names=FALSE)
```

#when reproductive investment is scaled by leaf area there is no significant difference by seed size - so increase in repro inv with seed size simply because species with bigger seeds are bigger plants  
```{r}
plot(scaled_repro_inv~seed_size,SummaryInd, col=col.spp(species),pch=16,log="xy")
points(scaled_repro_inv~seed_size,SummarySpp[["mean"]], cex=2, col=col.spp(species),pch=16,log="xy")
mod <-sma(scaled_repro_inv~seed_size,SummarySpp[["mean"]],log="xy",method="OLS")
summary(mod)
```

#Consider how accessory cost components shift in terms of costs per seed, and cost as fraction total reproductive investment

```{r}
par(mfrow=c(1,1), cex=1, omi=c(.1,.1,.1,.05), mai=c(.8,.8,.1,.1))

win.metafile("figs2/Accessory_Costs/Figure02_seedsize_vs_reproinv.wmf", height=5, width=5)
plot

par(mfrow=c(1,1), cex=1, omi=c(.1,.1,.1,.05), mai=c(.8,.8,.1,.1))
plot(seedset~seed_size,SummarySpp[["mean"]],log="xy",col=col.spp(species),pch=16)
mod <- lm(log10(seedset)~log10(seed_size),SummarySpp[["mean"]])
abline(mod$coeff[1],mod$coeff[2])
words.bottom.left.logxy(mod)
y.error.bars(SummarySpp[["mean"]]$seed_size,SummarySpp[["mean"]]$seedset,SummarySpp[["se"]]$seedset)

@@

#Table showing accessory cost values for the species
```{r}
acc_costs_spp <- dplyr::select(SummarySpp[["mean"]],species,seed_size,seedset,accessory_per_seed,prepollencosts,prepollen_all_per_seed,prepollen_failure_per_seed,postpollen_aborted_per_seed,packaging_dispersal_per_seed,prop_accessory_nonzero)
acc_costs_spp$species <- labels.spp.full()
acc_costs_spp <- format(acc_costs_spp, digit=2)

write.csv(acc_costs_spp, file="figs2/Accessory_Costs/Table02_AccCosts_by_Species.csv",row.names=FALSE)
```

#Figure 2 and Table 3. Scaling of accessory costs versus seed size, by species
```{r}
win.metafile("figs2/Accessory_Costs/Figure02_seed_size_correlations.wmf", height=5, width=15)
plot

@@

plot(prepollencosts~seed_size,SummarySpp[["mean"]],col=col.spp(species),pch=16,log="xy", yaxt="n",xaxt="n", ylim=c(.02,20000), xlim=c(.02,20000), ylab="required pre-pollination costs per seed (mg/seed)",xlab="",cex=1.5)
y.error.bars(SummarySpp[["mean"]]$seed_size,SummarySpp[["mean"]]$prepollencosts,SummarySpp[["se"]]$prepollencosts)
mod <- lm(log10(prepollencosts)~log10(seed_size),SummarySpp[["mean"]])
abline(coef(mod)[1],coef(mod)[2])
abline(coef(mod)[1],1,col="red")
axis(2, at=c(0.1,10,1000), labels=c(0.1,10,1000),cex.axis=1)
axis(1, at=c(0.1,10,1000), labels=c(0.1,10,1000),cex.axis=1)

plot(packaging_dispersal_per_seed~seed_size,SummarySpp[["mean"]],col=col.spp(species),pch=16,log="xy", yaxt="n",xaxt="n", ylim=c(.02,20000), xlim=c(.02,20000), ylab="packaging and dispersal costs per seed (mg/seed)",xlab="",cex=1.5)
y.error.bars(SummarySpp[["mean"]]$seed_size,SummarySpp[["mean"]]$packaging_dispersal_per_seed,SummarySpp[["se"]]$prepollencosts)
mod <- lm(log10(packaging_dispersal_per_seed)~log10(seed_size),SummarySpp[["mean"]])
abline(coef(mod)[1],coef(mod)[2])
abline(coef(mod)[1],1,col="red")
axis(2, at=c(0.1,10,1000), labels=c(0.1,10,1000),cex.axis=1)
axis(1, at=c(0.1,10,1000), labels=c(0.1,10,1000),cex.axis=1)
mtext(text="seed size (mg)",side=1,line=1.5,outer=TRUE)
dev.off()
```

```{r}
par(mfrow=c(2,3), cex=1, omi=c(.6,.1,.05,.05), mai=c(.1,.8,.05,.05)) 
options(scipen = 5)

@@

write.csv(acc_by_species, file="figs2/Accessory_Costs/Table03_acc_by_species.csv",row.names=FALSE)
```

#Table 4 and Figure 3.Per seed accessory costs are lower in individuals with higher seed count. 
###This effect is true across species, since larger-seeded species produce fewer seeds and have higher per seed accessory costs and true within species, since absolute accessory investment is a function of plant size not seedset
```{r}
win.metafile("figs2/Accessory_Costs/Figure_EXTRA_acc_cost_by_seed_count.wmf", height=5, width=5)
plot

@@

results
write.csv(results, file="figs2/Accessory_Costs/Table04a_seedcount_vs_accessory.csv",row.names=FALSE)

@@

#Table 4 extra - showing propagule investment instead of seed_count. Could also replace with seedset
```{r}
win.metafile("figs2/Accessory_Costs/Figure_acc_cost_by_seedset.wmf", height=9, width=9)
plot

@@

write.csv(results, file="figs2/Accessory_Costs/Table04b_seedset_vs_accessory.csv",row.names=FALSE)

@@

write.csv(results, file="figs2/Accessory_Costs/Table06a_prepollen_all_per_seed_vs_seedset.csv",row.names=FALSE)

@@

write.csv(results, file="figs2/Accessory_Costs/Table06b_postpollen_aborted_per_seed_vs_seedset.csv",row.names=FALSE)

@@

win.metafile("figs2/Accessory_Costs/Table07_dispcosts_vs_ovules.wmf", height=6, width=6)
plot

@@

write.csv(results, file="figs2/Accessory_Costs/Table06c_dispersal_vs_seedset.csv",row.names=FALSE)
```

#Higher RE, RA, plant size do not lead to lower accessory costs; age shows mixed results
##Not a single species shows a reduction in per seed accessory costs with increased plant size.Indeed, EPMI has higher pre-pollination accessory costs    (slope = 0.38, r2=0.19, p=0.019) and accessory costs due to aborted fruit and seeds (slope = 0.83, r2=0.268, p=0.019) in larger individuals, as larger individuals of this species have lower seedset (slope = -0.31, r2=0.198, p=0.018). GRSP also has a nearly significant decline in seedset in larger individuals (slope = -0.78, r2=0.298, p=0.067).
```{r}
win.metafile("figs2/Accessory_Costs/Figure_RE_RA_size_age_patterns.wmf", height=10, width=10)
plot

write.csv(results, file="figs2/Accessory_Costs/Table05a_accessory_vs_RE.csv",row.names=FALSE)

write.csv(results, file="figs2/Accessory_Costs/Table05c_accessory_vs_total_weight.csv",row.names=FALSE)

write.csv(results, file="figs2/Accessory_Costs/Table05d_accessory_vs_age.csv",row.names=FALSE)
dev.off()
```

```{r}
win.metafile("figs2/Accessory_Costs/Figure_seedcount_reprocount.wmf", height=5, width=5)
plot

dev.off()
```
#Considering the individual components of accessory costs vs. seedset - since seedset best predictor of accessory costs

#For each species the accessory cost components that cause shifts in total accessory costs is different and reflects both which components are largest and whichvary most with some factor - seedset, age, etc.
```{r}
xvar1 <- "packaging_dispersal_per_seed"
xvar2 <- "prepollencosts"
xvar3 <- "postpollen_aborted_per_seed"
xvar4 <- "prepollen_failure_per_seed"
yvar <- "accessory_per_seed"
data <- subset(SummaryInd,SummaryInd[[xvar1]]>0&SummaryInd[[xvar2]]>0&SummaryInd[[xvar3]]>0&SummaryInd[[xvar4]]>0&SummaryInd[[yvar]]>0&species!="BAER"&species!="COER"&species!="LEES")
data <- split(data, data$species)

@@

write.csv(species_coef, file="figs2/Accessory_Costs/Table08_accessory_components_coeff.csv",row.names=FALSE)

@@

write.csv(species_t, file="figs2/Accessory_Costs/Table08_accessory_components_t-value.csv",row.names=FALSE)

@@

write.csv(results, file="figs2/Accessory_Costs/Table_prop_tissues_inv_by_seed_size.csv",row.names=FALSE) 
```

@@

```{r}
win.metafile("figs2/Accessory_Costs/Figure_success_vs_failure.wmf", height=6, width=8)
plot

```{r}
ternary <- select(SummarySpp[["mean"]],species,seed_size,prepollencosts,packaging_dispersal_per_seed)
ternary$packaging_dispersal_per_seed[9] <- 0.1

write.csv(costs_per_seed, file="figs2/Accessory_Costs/Table08_costs_per_seed.csv",row.names=FALSE)

win.metafile("figs2/Accessory_Costs/seedcost_components.wmf", height=6, width=6)
plot

```{r}
win.metafile("figs2/Accessory_Costs/Ternary.wmf", height=5, width=5)
plot

>>>>

means, counts to include in manuscript
```{r}
count(SummaryInd)
count(subset(SummaryInd,seed_count>0))
count(subset(SummaryInd,repro_inv>0))
mean(subset(SummaryInd$prop_failure,SummaryInd$prop_failure>0))
mean(subset(SummaryInd$prop_accessory,SummaryInd$seed_count>0))
```

@@

write.csv(results, file="figs2/Accessory_Costs/Table01_correlations_with_repro_inv.csv",row.names=FALSE)
```

#Figure 2. Correlations of various reproductive investment categories against plant weight
```{r}
win.metafile("figs2/Accessory_Costs/Figure02_correlations_with_plant_weight.wmf", height=5, width=10)
plot

@@

#Table 3 showing accessory cost values for the species
```{r}
acc_costs_spp <- dplyr::select(SummarySpp[["mean"]],species,seed_size,accessory_per_seed,prop_accessory_nonzero,seedcosts,prop_success,
                               prepollencosts,packaging_dispersal_per_seed,failure_per_seed,prop_failure,prepollen_failure_per_seed,postpollen_aborted_per_seed, seedset)
acc_costs_spp$species <- labels.spp.full()
acc_costs_spp <- format(acc_costs_spp, digit=2)

write.csv(acc_costs_spp, file="figs2/Accessory_Costs/Table03_AccCosts_by_Species.csv",row.names=FALSE)
```

#Figure 3. Ternary diagram showing seedcosts
```{r}
ternary <- select(SummarySpp[["mean"]],species,seed_size,prepollencosts,packaging_dispersal_per_seed)
ternary$packaging_dispersal_per_seed[9] <- 0.1

write.csv(costs_per_seed, file="figs2/Accessory_Costs/Table00_in_text_costs_per_seed.csv",row.names=FALSE)

win.metafile("figs2/Accessory_Costs/FigureXX_seedcost_components.wmf", height=6, width=6)
plot

```{r}
win.metafile("figs2/Accessory_Costs/Figure03_ternary_diagram_seedcosts.wmf", height=5, width=5)
plot

@@

#Table 4 showing seedcost breakdown into components, by species
```{r}
prop_seedcosts_spp <- dplyr::select(SummarySpp[["mean"]],species,seedcosts,prop_prepollencosts,prop_dispersalcosts,prop_seedcosts)
prop_seedcosts_spp$species <- labels.spp.full()
prop_seedcosts_spp <- format(prop_seedcosts_spp, digit=2)
write.csv(prop_seedcosts_spp, file="figs2/Accessory_Costs/Table04_Prop_seedcosts_by_Species.csv",row.names=FALSE)
```

#Figure 4. Scaling of accessory costs versus seed size, by species
```{r}
win.metafile("figs2/Accessory_Costs/Figure04_seed_size_acc_cost_correlations.wmf", height=5, width=15)
plot

@@

plot(prepollencosts~seed_size,SummarySpp[["mean"]],col=col.spp(species),pch=16,log="xy", yaxt="n",xaxt="n", ylim=c(.02,20000), xlim=c(.02,20000), ylab="required pre-pollination costs per seed (mg/seed)",xlab="",cex=1.5)
#y.error.bars(SummarySpp[["mean"]]$seed_size,SummarySpp[["mean"]]$prepollencosts,SummarySpp[["se"]]$prepollencosts)
mod <- lm(log10(prepollencosts)~log10(seed_size),SummarySpp[["mean"]])
abline(coef(mod)[1],coef(mod)[2])
abline(coef(mod)[1],1,col="red")
axis(2, at=c(0.1,10,1000), labels=c(0.1,10,1000),cex.axis=1)
axis(1, at=c(0.1,10,1000), labels=c(0.1,10,1000),cex.axis=1)

plot(packaging_dispersal_per_seed~seed_size,SummarySpp[["mean"]],col=col.spp(species),pch=16,log="xy", yaxt="n",xaxt="n", ylim=c(.02,20000), xlim=c(.02,20000), ylab="packaging and dispersal costs per seed (mg/seed)",xlab="",cex=1.5)
#y.error.bars(SummarySpp[["mean"]]$seed_size,SummarySpp[["mean"]]$packaging_dispersal_per_seed,SummarySpp[["se"]]$prepollencosts)
mod <- lm(log10(packaging_dispersal_per_seed)~log10(seed_size),SummarySpp[["mean"]])
abline(coef(mod)[1],coef(mod)[2])
abline(coef(mod)[1],1,col="red")
axis(2, at=c(0.1,10,1000), labels=c(0.1,10,1000),cex.axis=1)
axis(1, at=c(0.1,10,1000), labels=c(0.1,10,1000),cex.axis=1)
mtext(text="seed size (mg)",side=1,line=1.5,outer=TRUE)
dev.off()
```

#Table 5. Slopes for various seed size by acc cost regressions
```{r}
par(mfrow=c(2,3), cex=1, omi=c(.6,.1,.05,.05), mai=c(.1,.8,.05,.05)) 
options(scipen = 5)

@@

write.csv(acc_by_species, file="figs2/Accessory_Costs/Table05_acc_cost_seed_size_slopes.csv",row.names=FALSE)
```

#Figure 5 seed count, scaled to plant size, decreases 1:1 with seed size, while bud count decreases less steeply; indication that seed count is what is actually "programmed in"
```{r}

win.metafile("figs2/Accessory_Costs/Figure05_cost_of_flowers_vs_seed_size.wmf", height=5, width=15)
plot

summary(mod2)
mod1 <- sma(scaled_seed_count~seed_size,data=subset(SummaryInd,scaled_seed_count!=0),log="xy",slope=-1,method="OLS")
mod2 <- sma(scaled_bud_count~seed_size,data=subset(SummaryInd,scaled_bud_count!=0),log="xy",slope=-1,method="OLS")
mod3 <- sma(scaled_failure_count~seed_size,data=subset(SummaryInd,scaled_failure_count!=0),log="xy",slope=-1,method="OLS")
sma_abline(mod1)
abline(mod2$coef[[1]][1,1],mod2$coef[[1]][2,1],col="blue")
abline(mod3$coef[[1]][1,1],mod3$coef[[1]][2,1],col="red")
legend("topright",legend=c("seed count","total bud count","failed flower count"),col=c("black","blue","red"),pch=c(16,21,8),bty="n")

#Table 6. Higher RE, RA, plant size do not lead to lower accessory costs; age shows mixed results
```{r}
win.metafile("figs2/Accessory_Costs/FigureXX_RE_RA_size_age_patterns.wmf", height=10, width=10)
plot

write.csv(results, file="figs2/Accessory_Costs/Table06a_accessory_vs_RE.csv",row.names=FALSE)

write.csv(results, file="figs2/Accessory_Costs/Table06b_accessory_vs_total_weight.csv",row.names=FALSE)

write.csv(results, file="figs2/Accessory_Costs/Table06c_accessory_vs_age.csv",row.names=FALSE)
dev.off()
```

###GOOD VISUALS - not currently in manuscript
```{r}
win.metafile("figs2/Accessory_Costs/FigureXX_seedcount_reprocount.wmf", height=5, width=5)
plot

```{r}
win.metafile("figs2/Accessory_Costs/FigureXX_success_vs_failure.wmf", height=6, width=8)
plot

#EXTRAS - not in manuscript
#Neither within nor across species, does the proportion of reproductive energy spent on "success" correlate with total reproductive investment
```{r}
par(mfrow=c(1,1), cex=1, omi=c(.1,.1,.1,.05), mai=c(.8,.8,.1,.05))

write.csv(results, file="figs2/Accessory_Costs/Table02_extra_propagule_invest_vs_repro_inv.csv",row.names=FALSE)
```

#when reproductive investment is scaled by leaf area there is no significant difference by seed size - so increase in repro inv with seed size simply because species with bigger seeds are bigger plants  
```{r}
plot(scaled_repro_inv~seed_size,SummaryInd, col=col.spp(species),pch=16,log="xy")
points(scaled_repro_inv~seed_size,SummarySpp[["mean"]], cex=2, col=col.spp(species),pch=16)
mod <-sma(scaled_repro_inv~seed_size,SummarySpp[["mean"]],log="xy",method="OLS")
summary(mod)
```

#Consider how accessory cost components shift in terms of costs per seed, and cost as fraction total reproductive investment
```{r}
par(mfrow=c(1,1), cex=1, omi=c(.1,.1,.1,.05), mai=c(.8,.8,.1,.1))

par(mfrow=c(1,1), cex=1, omi=c(.1,.1,.1,.05), mai=c(.8,.8,.1,.1))
plot(seedset~seed_size,SummarySpp[["mean"]],log="xy",col=col.spp(species),pch=16)
mod <- lm(log10(seedset)~log10(seed_size),SummarySpp[["mean"]])
abline(mod$coeff[1],mod$coeff[2])
words.bottom.left.logxy(mod)
y.error.bars(SummarySpp[["mean"]]$seed_size,SummarySpp[["mean"]]$seedset,SummarySpp[["se"]]$seedset)
```

#This effect is true across species, since larger-seeded species produce fewer seeds and have higher per seed accessory costs and true within species, since absolute accessory investment is a function of plant size not seedset
```{r}
win.metafile("figs2/Accessory_Costs/FigureXX_EXTRA_acc_cost_by_seed_count.wmf", height=5, width=5)
plot

@@

write.csv(results, file="figs2/Accessory_Costs/TableXX_seedcount_vs_accessory.csv",row.names=FALSE)

@@

#propagule investment instead of seed_count. Could also replace with seedset
```{r}
win.metafile("figs2/Accessory_Costs/FigureXX_acc_cost_by_seedset.wmf", height=9, width=9)
plot

@@

write.csv(results, file="figs2/Accessory_Costs/TableXX_seedset_vs_accessory.csv",row.names=FALSE)

@@

write.csv(results, file="figs2/Accessory_Costs/TableXX_prepollen_all_per_seed_vs_seedset.csv",row.names=FALSE)

@@

write.csv(results, file="figs2/Accessory_Costs/TableXX_postpollen_aborted_per_seed_vs_seedset.csv",row.names=FALSE)

@@

win.metafile("figs2/Accessory_Costs/FigureXX_dispcosts_vs_ovules.wmf", height=6, width=6)
plot

@@

write.csv(results, file="figs2/Accessory_Costs/TableXX_dispersal_vs_seedset.csv",row.names=FALSE)
```

#Considering the individual components of accessory costs vs. seedset - since seedset best predictor of accessory costs

#For each species the accessory cost components that cause shifts in total accessory costs is different and reflects both which components are largest and whichvary most with some factor - seedset, age, etc.
#CODE CURRENTLY DOESN"T WORK
```{r}
xvar1 <- "packaging_dispersal_per_seed"
xvar2 <- "prepollencosts"
xvar3 <- "postpollen_aborted_per_seed"
xvar4 <- "prepollen_failure_per_seed"
yvar <- "accessory_per_seed"
data <- subset(SummaryInd,SummaryInd[[xvar1]]>0&SummaryInd[[xvar2]]>0&SummaryInd[[xvar3]]>0&SummaryInd[[xvar4]]>0&SummaryInd[[yvar]]>0&species!="BAER"&species!="COER"&species!="LEES")
data <- split(data, data$species)

@@

write.csv(species_coef, file="figs2/Accessory_Costs/TableXX_accessory_components_coeff.csv",row.names=FALSE)

@@

write.csv(species_t, file="figs2/Accessory_Costs/TableXX_accessory_components_t-value.csv",row.names=FALSE)

@@

write.csv(results, file="figs2/Accessory_Costs/TableXX_prop_tissues_inv_by_seed_size.csv",row.names=FALSE) 
```

@@
